Missing the Point$\colon$ Non-Convergence in Multiple Imputation Procedures

### 2020-04-01 3.Thesis/1.SimulationStudy/Writeup/manuscript.Rmd
Read entire research report and put comments in it for thesis manuscript. Ran 5,25,50,75,and 95% miss with 100 sims each.

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
load("../Results/results.Rdata")
source("../2.CreateFigures.R")

@@

At some point, any scientist conducting statistical analyses will run into a missing data problem [@alli02]. Missingness is problematic because statistical inference cannot be performed on incomplete data without employing \emph{ad hoc} solutions (e.g., list-wise deletion), which may yield wildly invalid results [@buur18]. A popular answer to the ubiquitous problem of missing information is to use the framework of multiple imputation (MI), proposed by @rubin87. For the theoretical foundation of MI, see @rubin87. For an accessible and comprehensive introduction to MI from an applied perspective, see e.g. @buur18. With MI, missing datapoints are 'imputed' (i.e., filled in) via an *imputation model*. The imputation model can be a joint model for all variables, or several univariate conditional distributions (fully conditional specification). FCS is more flexible and "avoids the problem of specifying an appropriate joint imputation distribution and replaces this by the selection of appropriate univariate conditional distributions". With FCS, an iterative algorithmic procedure is used to ... multiple imputation by chained equations ... "The values for imputing the missing data are drawn from the posterior predictive distribution of the missing data conditional on the observed data."  There is no scientific consensus on the convergence properties of FCS algorithms [@taka17]. Some default techniques in `mice` might not yield converged states at all [@murr18]. Therefore, algorithmic convergence should be monitored carefully.

@@

Current practice is visually inspecting imputation chains for signs of non-convergence, but may be challenging to the untrained eye [@buur18, $\S$ 6.5.2]. Therefore diagnostic evaluation of convergence would be preferred. However, not known whether MCMC diagnostics are appropriate for MICE ... The application of convergence diagnostics to MI methods has not been systematically studied [@buur18]. How can non-convergence be diagnosed? How many iterations are sufficient/required? What are the effects of non-convergence on estimates, predictions and inference? What is an appropriate non-convergence diagnostic? Are the default number of iterations sufficient? Defaults in: SPSS = 10 [(link)](https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html), Mplus = 100?? [(link)](https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748), Stata = 1; [(link)](https://www.stata.com/manuals13/mi.pdf, p. 139), Amelia = NA, because resampling, not convergence [(link)](https://cran.r-project.org/web/packages/Amelia/Amelia.pdf), en MI = 30 [(link)](https://cran.r-project.org/web/packages/mi/mi.pdf).

@@

Let $y$ denote an $n \times p$ matrix containing the data values on $p$ variables for all $n$ units in a sample. The data value of unit $i$ ($i = 1, \dots, n$) on variable $j$ ($j = 1, \dots, p$) may be either observed or missing. The collection of observed data values in $y$ is denoted by $y_{obs}$; the missing part of $y$ is referred to as $y_{mis}$. For each datapoint in $y_{mis}$ $M \times T$ times plausible values are sampled, where $M$ is the number of imputations ($m = 1, \dots, M$) and $T$ is the number of iterations ($t = 1, \dots, T$). The collection of samples between the initial value (at $t=1$) and the imputed value (at $t=T$) will be referred to as an 'imputation chain'.

@@

In iterative algorithmic procedures (Markov chain Monte Carlo methods; e.g., `mice` algorithms or Gibbs samplers) non-convergence may be present as non-stationarity within chains (i.e., trending), or as slow mixing between chains (i.e., no intermingling). (**add example/guidelines visual inspection here?**)

The stationarity component of convergence may be evaluated with autocorrelation [$AC$; @scha97; @gelm13], numeric standard error [or `MC error'; @gewe92], and Raftery and Lewis's [-@raft91] procedure to determine the effect of trending within chains.

The mixing component can be assessed with the potential scale reduction factor $\widehat{R}$ [a.k.a. `Gelman-Rubin statistic'; @gelm92]. With an adapted version of $\widehat{R}$, proposed by @veht19, we might also evaluate the stationarity component of convergence. This would make $\widehat{R}$ a general convergence diagnostic. The application of $\widehat{R}$ to assess stationarity has not been thoroughly investigated. Therefore, this study employs both $\widehat{R}$ and autocorrelation to investigate convergence, as recommended by [@cowl96, p. 898].

@@

\begin{align*}
B&=\frac{T}{M-1} \sum_{m=1}^{M}\left(\bar{\theta}^{(\cdot m)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \text { where } \bar{\theta}^{(\cdot m)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t m)}, \quad \bar{\theta}^{(\cdot \cdot)}=\frac{1}{M} \sum_{m=1}^{M} \bar{\theta}^{(\cdot m)}. 
\end{align*}
<!-- %\text{ \cite[p.~5]{veht19}}  -->

@@

This study evaluates whether $\widehat{R}$ and $AC$ could diagnose convergence of multiple imputation algorithms. We assess the performance of the two convergence diagnostics against the recommended evaluation criteria for MI methods [i.e., average bias, average confidence interval width, and empirical coverage rate across simulations; @buur18, $\S$ 2.5.2]. That is, there is no baseline measure available to evaluate performance against.The current practice of visually inspecting imputation chains for signs of non-mixing or non-stationarity can only diagnose severely pathological cases of non-convergence. We could also look at distributional characteristics, and plausibility of imputed values, see @vinknd. For now, this is outside of the scope of this study.

Based on an empirical finding [@lace07], we hypothesize that $\widehat{R}$ will over-estimate non-convergence of MI algorithms. The threshold of $\widehat{R} < 1.01$ will then be too stringent for diagnosing convergence. This over-estimation may, however, be diminished because $\widehat{R}$ can falsely diagnose convergence if initial values of the algorithm are not appropriately over-dispersed [@broo98, p. 437]. In `mice`, initial values are chosen randomly from the observed data. Therefore, we cannot be certain that the initial values are over-dispersed. We expect this to have little effect on the hypothesized performance of $\widehat{R}$. No hypothesis was formulated about the performance of $AC$ as convergence diagnostic.

@@

A finite population of $N=1000$ is simulated to solve a multiple linear regression problem, where dependent variable $Y$ is regressed on independent variables $X_1$, $X_2$ and $X_3$:

@@

<!-- From the imputation step, we extract chain means and chain variances across iterations ($\theta$ at $t = 1,2,\dots,T$).   -->
<!-- We compute convergence diagnostics separately for both $\theta$s, and for each of the four variables in the dataset. We report the maximum (absolute) value as the 'worst linear predictor' of convergence.  -->
$\widehat{R}$ is computed by implementing Vehtari et al.'s [-@veht19] recommendations. $AC$ is computed with function `stats::acf()`.

@@

We compute bias as the difference between $\bar{Q}$ and $Q$. Confidence interval width (CIW) is defined as the difference between the lower and upper bound of the 95\% confidence interval (CI95\%) around $\bar{Q}$. We compute the CI95\% bounds as

@@

**Add more info about figure legends and axes.**

@@

## Convergence Diagnostics

It is apparent that there is a relation between the number of iterations per simulation condition ($T$) and the convergence diagnostics. Generally speaking, conditions with longer imputation chains (higher $T$) coincide with less signs of non-convergence ($\widehat{R}$-values approach one, and $AC$-values approach zero). We see more or less equivalent trends for chain means versus chain variances. We, therefore, only discuss the $\widehat{R}$ and $AC$-values of chain means.

Figure 2A shows that $\widehat{R}$-values generally decrease with increasing imputation chain lengths. The decline stabilizes somewhere between the simulation conditions $T=30$ and $T=50$. The downward trend is most pronounced for $T=3$, and between $T = 5$ and $T = 10$. In the intervening conditions ($3 \leq T \leq 5$), however, we observe a steep increase in $\widehat{R}$-values. This increase implies that non-convergence is under-estimated, and convergence should not be diagnosed at this point. Based on the conventional threshold $\widehat{R} < 1.2$, we would falsely diagnose convergence at $T=3$. According to the widely used threshold $\widehat{R} < 1.1$, convergence would be diagnosed for conditions where $T>9$. If we use the recently recommended threshold $\widehat{R} < 1.01$, we would conclude that convergence is reached in none of the simulation conditions.

The $AC$-values displayed in Figure 2B are almost uniformly increasing as a function of $T$. The only $AC$-value that deviates from this observation is for $T=2$. The lowest $AC$-value is obtained for $T=3$. %At $T=5$, the $AC$-value reaches the level observed at $T=2$. 
The gradual increase plateaus between $T=10$ and $T=30$. $AC$-values in conditions where $T>70$ are indifferentiable from zero, indicating stationarity. We see an initial decrease between $T=2$ and $T=3$. Simulation conditions where $T>5$ have $AC$-values greater than $T=2$.

According to this diagnostic, none of the simulation conditions show signs of non-convergence, since we only observe negative or zero $AC$-values. The negative $AC$-values, however, indicate increasing divergence between subsequent samples in conditions where $T<4$. It may not be wise to terminating the algorithm at that point, but after `recovering' from this dip in $AC$-values. This would imply $T=6$ as the minimal number of iterations.

@@

We use average bias, average confidence interval width, and coverage rate as performance measures to evaluate $\widehat{R}$ and $AC$. We make the general observation that the simulation diagnostics behave as theorized for most simulation conditions (bias around zero, stable confidence interval widths, nominal coverage rate at 95\%).

Figure 3A shows that bias is fairly stable across simulation conditions. The condition $T=1$ clearly deviates from this trend with a negative bias. The bias for $T=2$ is below average, but within the range of fluctuations. Conditions where $T>3$ can be diagnosed as unbiased, but the average bias across iterations (as shown by a flat Loess line) only reaches stability at $T=20$.

CIW is only clearly divergent in the simulation condition $T=1$, see Figure 3B. However, under-estimating the variance of $\bar{Q}$, which is the case for $T=3$, may yield spurious inferences. Only conditions where $T>3$ are therefore considered sufficient. These conditions have similar CIWs, but across iterations stability is not established (i.e., the Loess line is never completely flat within the 100 simulation conditions). %After about twenty-five iterations the average CIW across all simulation conditions and repetitions is reached 0.927.

The empirical coverage rate across repetitions seems more or less stable for conditions where $T>1$, see Figure 3C. We see some over-coverage at $T=2$, but that is better than under-estimating the variance of $\bar{Q}$. On average, the coverage rate is somewhat higher than the expected nominal coverage of 95\% [@neym34], namely 95\%. Similar to the CIWs, there is some trending across iterations (i.e., the Loess line is never flat).

@@

This note shows that convergence diagnostics $\widehat{R}$ and $AC$ may diagnose convergence of multiple imputation algorithms, but their performance differs from conventional applications to iterative algorithmic procedures.   
$\widehat{R}$ and autocorrelation indicate that algorithmic convergence may only be reached after twenty or even forty iterations, while unbiased, confidence valid estimates estimates may be obtained with as little as four iterations. These results are in agreement with the simulation hypothesis: $\widehat{R}$ over-estimates the severity of non-convergence when applied to MI procedures. %This may be due to the quantity of scientific interest chosen. More `complicated' $Q$s (e.g., higher order effects or variance components) might show bias, under- or over-coverage at higher $T$.

According to this simulation study, the recently proposed threshold of $\widehat{R}<1.01$ may be too stringent for MI algorithms. Under the relatively easy missing data problem of the current study, the threshold was not reached. The other extreme of the $\widehat{R}$-thresholds, the conventionally acceptable $\widehat{R} <1.2$, may be too lenient for MI procedures. Applying this threshold to the current data, lead to falsely diagnosing convergence at $T = 3$. It appears that the widely used threshold of $\widehat{R} < 1.1$ suits MI algorithms the best. We might, however, also formulate a new threshold, specifically for the evaluation of MI algorithms. The current study suggests that $\widehat{R} < 1.05$ may be implemented, since that is the level at which the $\widehat{R}$ stabilize (around $T = 20$).

The negative $AC$-values obtained in this study show no threat of non-stationarity. However, initial dip in $AC$-values may have implications for the default number of iterations in `mice` (`maxit = 5`). Terminating the algorithm at $T=5$ may not be the most appropriate, since this lead to the worst convergence, as indicated by $\widehat{R}$ and $AC$. Under the current specifications, $T>20$ would be more appropriate.

The observed dip in AC implies that default maxit value of five iterations is the worst possible number of iterations. Moreover, the results of this study imply that assessing the stationarity component of convergence with $AC$ might be redundant, since high $AC$-values are implausible in MI procedures. That is, the randomness induced by the MI algorithm effectively mitigates the risk of dependency within chains. $AC$ would thus not be informative of the convergence of MI algorithms.

Since this study only considers only an MCAR missingness mechanism, results may not be extrapolated to other missing data problems. Proper performance of the convergence diagnostics under MCAR is necessary but not sufficient to demonstrate appropriateness of $\widehat{R}$ and $AC$ as convergence diagnostics. This is just a proof of concept.

Further research is needed to investigate their performance under clear violation of convergence, e.g. dependency between predictors (predictors with very high correlations). Until then, we have only shown that the convergence diagnostics can diagnose non-convergence of MI algorithms that trend towards a converged state. Also for future research, implement Wernicke diagnostic, and look at developing a convergence diagnostic for substantive models, and implement a Wald test for $AC = 0$.

>>>>

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 5, echo = FALSE, fig.align = "center") 
load("../Results/results.Rdata")
source("../2.CreateFigures.R")

@@

At some point, any scientist conducting statistical analyses will run into a missing data problem [@alli02]. Missingness is problematic because statistical inference cannot be performed on incomplete data without employing \emph{ad hoc} solutions (e.g., list-wise deletion), which may yield wildly invalid results [@buur18]. A popular answer to the ubiquitous problem of missing information is to use the framework of multiple imputation (MI), proposed by @rubin87. **(Call it mice already, then explain mi and ce sides of mice)** For the theoretical foundation of MI, see @rubin87. For an accessible and comprehensive introduction to MI from an applied perspective, see e.g. @buur18. With MI, missing datapoints are 'imputed' (i.e., filled in) via an *imputation model* (**term not required to understand this paper, so circumvent**) . (**Move this to mi vs. ce part in intro:** The imputation model can be a joint model for all variables, or several univariate conditional distributions (fully conditional specification). FCS is more flexible and "avoids the problem of specifying an appropriate joint imputation distribution and replaces this by the selection of appropriate univariate conditional distributions".)**

**(Rewrite: With FCS, an iterative algorithmic procedure is used to ... multiple imputation by chained equations ... "The values for imputing the missing data are drawn from the posterior predictive distribution of the missing data conditional on the observed data.")**

There is no scientific consensus on the convergence properties of FCS algorithms [@taka17]. Some default techniques in `mice` might not yield converged states at all [@murr18]. Therefore, algorithmic convergence should be monitored carefully.

@@

Current practice is visually inspecting imputation chains for signs of non-convergence, but may be challenging to the untrained eye [@buur18, $\S$ 6.5.2]. Therefore diagnostic evaluation of convergence would be preferred. However, not known whether MCMC diagnostics are appropriate for MICE ... The application of convergence diagnostics to MI methods has not been systematically studied [@buur18]. How can non-convergence be diagnosed? How many iterations are sufficient/required? What are the effects of non-convergence on estimates, predictions and inference? What is an appropriate non-convergence diagnostic? Are the default number of iterations sufficient? Defaults in: SPSS = 10 [(link)](https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html), Mplus = 100?? [(link)](https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748), Stata = 10 [(link)](https://www.stata.com/manuals13/mi.pdf, p. 139), Amelia = NA, because resampling, not convergence [(link)](https://cran.r-project.org/web/packages/Amelia/Amelia.pdf), en MI = 30 [(link)](https://cran.r-project.org/web/packages/mi/mi.pdf).

@@

Let $y$ denote an $n \times p$ matrix containing the data values on $p$ variables for all $n$ units in a sample. The data value of unit $i$ ($i = 1, \dots, n$) on variable $j$ ($j = 1, \dots, p$) may be either observed or missing. The collection of observed data values in $y$ is denoted by $y_{obs}$; the missing part of $y$ is referred to as $y_{mis}$. **(This is not only notation, but theory too: For each datapoint in $y_{mis}$ $M \times T$ times plausible values are sampled, where $M$ is the number of imputations ($m = 1, \dots, M$) and $T$ is the number of iterations ($t = 1, \dots, T$).)** The collection of samples between the initial value (at $t=1$) and the imputed value (at $t=T$) will be referred to as an 'imputation chain'.

@@

**(Start with current guidelines, then continue with possible quant measures of the same things from MCMC lit.)** In iterative algorithmic procedures (Markov chain Monte Carlo methods; e.g., `mice` algorithms or Gibbs samplers) non-convergence may be present as non-stationarity within chains (i.e., trending), or as slow mixing between chains (i.e., no intermingling (**source??**)). (**add example/guidelines visual inspection here?**)

The stationarity component of convergence may be evaluated with autocorrelation [$AC$; @scha97; @gelm13], numeric standard error [or `MC error'; @gewe92], and Raftery and Lewis's [-@raft91] procedure to determine (**the chain length required to cancel out trending**) the effect of trending within chains.

The mixing component can be assessed with the potential scale reduction factor $\widehat{R}$ [a.k.a. `Gelman-Rubin statistic'; @gelm92]. With an adapted version of $\widehat{R}$, proposed by @veht19, the stationarity component of convergence might be evaluated as well. This would make $\widehat{R}$ a general convergence diagnostic. The application of $\widehat{R}$ to assess stationarity has not been thoroughly investigated. Therefore, this study employs both $\widehat{R}$ and autocorrelation to investigate convergence, as recommended by [@cowl96, p. 898].

@@

\begin{align*}
B&=\frac{T}{M-1} \sum_{m=1}^{M}\left(\bar{\theta}^{(\cdot m)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \text { where } \bar{\theta}^{(\cdot m)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t m)} \text{, } \bar{\theta}^{(\cdot \cdot)}=\frac{1}{M} \sum_{m=1}^{M} \bar{\theta}^{(\cdot m)}. 
\end{align*}
<!-- %\text{ \cite[p.~5]{veht19}}  -->

@@

This study evaluates whether $\widehat{R}$ and $AC$ could diagnose convergence of multiple imputation algorithms. We assess the performance of the two convergence diagnostics against the recommended evaluation criteria for MI methods [i.e., average bias, average confidence interval width, and empirical coverage rate across simulations; @buur18, $\S$ 2.5.2]. That is, there is no baseline measure available to evaluate performance against. **(Move above: The current practice of visually inspecting imputation chains for signs of non-mixing or non-stationarity can only diagnose severely pathological cases of non-convergence.)**
<!-- We could also look at distributional characteristics, and plausibility of imputed values, see @vinknd. For now, this is outside of the scope of this study. -->

Based on an empirical finding [@lace07], we hypothesize that $\widehat{R}$ will over-estimate non-convergence of MI algorithms. The threshold of $\widehat{R} < 1.01$ will then be too stringent for diagnosing convergence. This over-estimation may, however, be diminished because $\widehat{R}$ can falsely diagnose convergence if initial values of the algorithm are not appropriately over-dispersed [@broo98, p. 437]. In `mice`, initial values are chosen randomly from the observed data. Therefore, we cannot be certain that the initial values are over-dispersed. We expect this to have little effect on the hypothesized performance of $\widehat{R}$. **(Add actual hypothesis.)** No hypothesis was formulated about the performance of $AC$ as convergence diagnostic.

@@

A finite population of $N=1000$ is simulated to solve a multiple linear regression problem, where dependent variable $Y$ is regressed on independent variables $X_1$, $X_2$ and $X_3$ (**check notation with betas here, and add what the quantity/-ies of scientific interest is/are!**):

@@

<!-- From the imputation step, we extract chain means and chain variances across iterations ($\theta$ at $t = 1,2,\dots,T$).   -->
<!-- We compute convergence diagnostics separately for both $\theta$s, and for each of the four variables in the dataset. We report the maximum (absolute) value as the 'worst linear predictor' of convergence.  -->
$\widehat{R}$ (**of what?? i.e., chain means and chain variances of each variable**) is computed by implementing Vehtari et al.'s [-@veht19] recommendations. $AC$ (**of what? and how is AC aggregated across imputations? i.e., mean in latest sim**) is computed with function `stats::acf()`.

@@

We compute bias as the difference between $Q$ and $\bar{Q}$. Confidence interval width (CIW) is defined as the difference between the lower and upper bound of the 95\% confidence interval (CI95\%) around $\bar{Q}$. We compute the CI95\% bounds as

@@

**(Add more info about figure legends and axes.)**

@@

## Convergence Diagnostics (pair convergnce diagnostics with respective estimates, not split between convergence and simulation diagnostics)

**(Add different percentages missingness??)**

It is apparent that there is a relation between the number of iterations per simulation condition ($T$) and the convergence diagnostics. Generally speaking, conditions with longer imputation chains (higher $T$) coincide with less signs of non-convergence ($\widehat{R}$-values approach one, and $AC$-values approach zero). 
<!-- We see more or less equivalent trends for chain means versus chain variances. We, therefore, only discuss the $\widehat{R}$ and $AC$-values of chain means.  -->

Figure 2A shows that $\widehat{R}$-values generally decrease with increasing imputation chain lengths. The decline stabilizes somewhere between the simulation conditions $T=30$ and $T=50$. The downward trend is most pronounced for $T=3$, and between $T = 5$ and $T = 10$. In the intervening conditions ($3 \leq T \leq 5$), however, we observe a steep increase in $\widehat{R}$-values. This increase implies that non-convergence is under-estimated, and convergence should not be diagnosed at this point (**is that true???**). Based on the conventional threshold $\widehat{R} < 1.2$, we would falsely diagnose convergence at $T=3$. According to the widely used threshold $\widehat{R} < 1.1$, convergence would be diagnosed for conditions where $T>9$. If we use the recently recommended threshold $\widehat{R} < 1.01$, we would conclude that convergence is reached in none of the simulation conditions.

The $AC$-values displayed in Figure 2B are almost uniformly increasing as a function of $T$. The only $AC$-value that deviates from this observation is for $T=2$. The lowest $AC$-value is obtained for $T=3$. %At $T=5$, the $AC$-value reaches the level observed at $T=2$. 
The gradual increase plateaus between $T=10$ and $T=30$. $AC$-values in conditions where $T>70$ are indifferentiable from zero, indicating stationarity. We see an initial decrease between $T=2$ and $T=3$. Simulation conditions where $T>5$ have $AC$-values greater than $T=2$ (**weird wording**).

According to this diagnostic, none of the simulation conditions show signs of non-convergence, since we only observe negative or zero $AC$-values. The negative $AC$-values, however, indicate increasing divergence (**move above!**) between subsequent samples in conditions where $T<4$. It may not be wise (**why? and how would we know? e.g., wider distrubutions? didn't seem like it...**) to terminating the algorithm at that point, but after `recovering' from this dip in $AC$-values. This would imply $T=6$ as the minimal number of iterations.

@@

We use average bias, average confidence interval width, and coverage rate as performance measures to evaluate $\widehat{R}$ and $AC$ (**not the goal! we want to show the effects of non-conv, not just see how to diagnose it.**). We make the general observation that the simulation diagnostics behave as theorized for most simulation conditions (bias around zero, stable confidence interval widths, nominal coverage rate at 95\%).

Figure 3A shows that bias is fairly stable across simulation conditions. The condition $T=1$ clearly deviates from this trend with a negative bias. The bias for $T=2$ is below average, but within the range of fluctuations. Conditions where $T>3$ can be diagnosed as unbiased, but the average bias across iterations (as shown by a flat Loess line) only reaches stability at $T=20$ (**Loess might not be relevant, check and remove accordingly!**).

CIW is only clearly divergent in the simulation condition $T=1$, see Figure 3B. However, under-estimating the variance of $\bar{Q}$, which is the case for $T=3$, may yield spurious inferences (**move above**). Only conditions where $T>3$ are therefore considered sufficient. (**Possibly irrelevant Loess: These conditions have similar CIWs, but across iterations stability is not established (i.e., the Loess line is never completely flat within the 100 simulation conditions).**)

The empirical coverage rate across repetitions seems more or less stable for conditions where $T>1$, see Figure 3C. We see some over-coverage at $T=2$, but that is better than under-estimating the variance of $\bar{Q}$. On average, the coverage rate is somewhat higher than the expected nominal coverage of 95\% [@neym34] (**not anymore??**). (**Loess: Similar to the CIWs, there is some trending across iterations (i.e., the Loess line is never flat)**).

@@

This note shows that convergence diagnostics $\widehat{R}$ and $AC$ may diagnose convergence of multiple imputation algorithms, but their performance differs from conventional applications to iterative algorithmic procedures. (**nope! it shows that MICE can lead to correct outcomes when they have not converged accroding to two common conv diags. This may be due to the measures (e.g., assumption of overdisp) or due to the Qs (lm reg coeff, not higher dimensional/more complex RQs). Add what %miss has to do with it.**)  
$\widehat{R}$ and autocorrelation indicate that algorithmic convergence may only be reached after twenty or even forty iterations, while unbiased, confidence valid estimates estimates may be obtained with as little as four iterations. These results are in agreement with the simulation hypothesis: $\widehat{R}$ over-estimates the severity of non-convergence when applied to MI procedures. %This may be due to the quantity of scientific interest chosen. More `complicated' $Q$s (e.g., higher order effects or variance components) might show bias, under- or over-coverage at higher $T$.

According to this simulation study, the recently proposed threshold of $\widehat{R}<1.01$ may be too stringent for MI algorithms. (**This is only one of the goals: to give applied researchers a diagnostic to indicate that they should keep iterating. The other is the default in mice and other software packages, and yet another is ... i forgot**) Under the relatively easy missing data problem of the current study, the threshold was not reached. The other extreme of the $\widehat{R}$-thresholds, the conventionally acceptable $\widehat{R} <1.2$, may be too lenient for MI procedures. Applying this threshold to the current data, lead to falsely diagnosing convergence at $T = 3$ (**because it goes up after, not because it is not converged enough**). It appears that the widely used threshold of $\widehat{R} < 1.1$ suits MI algorithms the best. We might, however, also formulate a new threshold, specifically for the evaluation of MI algorithms. The current study suggests that $\widehat{R} < 1.05$ may be implemented, since that is the level at which the $\widehat{R}$ stabilize (around $T = 20$) (**Not necessary for this Q, but maybe for more complicated Qs**).

The negative $AC$-values obtained in this study show no threat of non-stationarity. However, initial dip in $AC$-values may have implications for the default number of iterations in `mice` (`maxit = 5`). Terminating the algorithm at $T=5$ may not be the most appropriate, since this lead to the worst convergence (**nope, only for Rh, not AC**), as indicated by $\widehat{R}$ and $AC$. Under the current specifications, $T>20$ would be more appropriate.

The observed dip in AC implies that default maxit value of five iterations is the worst possible number of iterations. Moreover, the results of this study imply that assessing the stationarity component of convergence with $AC$ might be redundant, since high $AC$-values are implausible in MI procedures. That is, the randomness induced by the MI algorithm effectively mitigates the risk of dependency within chains (**add above!!**). $AC$ would thus not be informative of the convergence of MI algorithms.

(**Add above, not here because it was a known limitation: Since this study only considers only an MCAR missingness mechanism, results may not be extrapolated to other missing data problems. Proper performance of the convergence diagnostics under MCAR is necessary but not sufficient to demonstrate appropriateness of $\widehat{R}$ and $AC$ as convergence diagnostics. This is just a proof of concept.**)

Further research is needed to investigate their performance under clear violation of convergence, e.g. dependency between predictors (predictors with very high correlations). Until then, we have only shown that the convergence diagnostics can diagnose non-convergence of MI algorithms that trend towards a converged state. Also for future research, look at developing a convergence diagnostic for substantive models, and implement a Wald test for $AC = 0$.
